---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# tjmisc 

<!-- badges: start -->
[![R-CMD-check](https://github.com/tjmahr/tjmisc/workflows/R-CMD-check/badge.svg)](https://github.com/tjmahr/tjmisc/actions)
<!-- badges: end -->

The goal of tjmisc is to gather miscellaneous helper functions, mostly for use 
in [my dissertation](https://github.com/tjmahr/dissertation). 

Apologies in advance. I think "misc" packages are kind of bad because packages
should be focused on specific problems: for example, my helper packages for
[working on polynomials](https://github.com/tjmahr/polypoly),
[printing numbers](https://github.com/tjmahr/printy) or 
[tidying MCMC samples](https://github.com/tjmahr/tristan). Having modular code
snapping together like Lego blocks is better than a grab-bag of functions, it's
true, but using `library(helpers)` is much, much better than using
`source("helpers.R")`. So here we are... in the grab-bag.


## Installation

You can install the tjmisc from github with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tjmahr/tjmisc")
```

## Examples

### Sample groups of data

`sample_n_of()` is like dplyr's `sample_n()` but it samples groups. 

```{r example}
library(dplyr, warn.conflicts = FALSE)
library(tjmisc)
set.seed(11022017)

data <- tibble::tibble(
  day = 1:10 %>% rep(10) %>% sort(),
  id  = 1:10 %>% rep(10),
  block = letters[1:5] %>% rep(10) %>% sort() %>% rep(2),
  value = rnorm(100) %>% round(2)
)

# data from 3 days
sample_n_of(data, 3, day)

# data from 1 id
sample_n_of(data, 1, id)

# data from 2 block-id pairs
sample_n_of(data, 2, block, id)
```

### Tidy quantiles

`tidy_quantile()` returns a dataframe with quantiles for a given variable. I
like to use it to select values for plotting model predictions.

```{r}
penguins <- palmerpenguins::penguins

penguins %>% 
  tidy_quantile(bill_length_mm)

penguins %>% 
  group_by(species) %>% 
  tidy_quantile(bill_length_mm)
```

### Tidy correlations

`tidy_correlation()` calculates correlations between pairs of selected dataframe
columns. It accepts `dplyr::select()` selection semantics, and it respects
grouped dataframes.

```{r}
penguins %>% 
  tidy_correlation(bill_length_mm, bill_depth_mm, flipper_length_mm)

penguins %>%
  dplyr::group_by(species) %>%
  tidy_correlation(dplyr::ends_with("mm"))
```

### Pairwise comparisons

`compare_pairs()` compares all pairs of values among levels of a categorical
variable. Hmmm, that sounds confusing. Here's an example. We compute the
difference in average score between each pair of workers.

```{r}
to_compare <- nlme::Machines %>%
  group_by(Worker) %>%
  summarise(avg_score = mean(score)) %>%
  print()

to_compare %>%
  compare_pairs(Worker, avg_score) %>%
  rename(difference = value) %>%
  mutate(
    across(where(is.numeric), round, 1)
  )
```

### Plotting a matrix

`ggmatplot()` plots the columns of a matrix as individual lines, much like 
`matplot()` in base R.

Here we plot a spline basis matrix for penguin bill length. By default it plots
the columns with unique row number as the x-axis.

```{r matplot, fig.width = 4, fig.height = 2.5, fig.retina = 2}
# Create a 10-column natural spline bases
sorted_lengths <- sort(penguins$bill_length_mm)
length_ns <- splines::ns(sorted_lengths, df = 10)
ggmatplot(length_ns)
```

Alternatively, you can supply a column number and make it the *x* axis. In this
example, we bind on the original data and use it as the *x*-axis column.
This makes the lines much smoother because the spline basis was built on the
bill lengths, not on row numbers.

```{r matplot1, fig.width = 4, fig.height = 2.5, fig.retina = 2}
ggmatplot(cbind(sorted_lengths, length_ns), x_axis_column = 1)
```

By default, duplicated rows are removed. We can choose to keep them. The little
flat steps along the curve are the repeated rows. We can also change the number
of colors to use. The package also provides `annotate_label_grey()` 
for making labels on ggplot2's default grey background.

```{r matplot2, fig.width = 4, fig.height = 2.5, fig.retina = 2}
ggmatplot(length_ns, unique_rows = FALSE, n_colors = 1) + 
  annotate_label_grey("splines!", 20, .65, size = 5)
```


### Et cetera

`ggpreview()` is like ggplot2's `ggsave()` but it saves an image to a temporary
file and then opens it in the system viewer. If you've ever found yourself in
a loop of saving a plot, leaving RStudio to doubleclick the file, sighing, going
back to RStudio, tweaking the height or width or plot theme, ever so slowly
spiraling in on your desired plot, then `ggpreview()` is for you.

`seq_along_rows()` saves a few keystrokes in for-loops that iterate over
dataframe rows.

```{r}
cars %>% head(5) %>% seq_along_rows()
cars %>% head(0) %>% seq_along_rows()
```

`is_same_as_last` and `replace_if_same_as_last()` are helpers for formatting
tables. I use them to replace repeating values in a text column with blanks.

```{r}
mtcars %>% 
  tibble::rownames_to_column("name") %>% 
  slice(1:10) %>% 
  select(cyl, name, mpg) %>% 
  arrange(cyl, mpg) %>% 
  mutate_at(c("cyl"), replace_if_same_as_last, "") %>% 
  knitr::kable()
```

`fct_add_counts()` adds counts to a factor's labels.

```{r}
# Create a factor with some random counts
set.seed(20190124)
random_penguins <- penguins %>% 
  dplyr::sample_n(250, replace = TRUE)

table(random_penguins$species)

# Updated factors
random_penguins$species %>% levels()
random_penguins$species %>% fct_add_counts() %>% levels()
```

You can tweak the format for the first label. I like to use this for plotting by
stating the unit next to the first count. 

```{r}
random_penguins$species %>% 
  fct_add_counts(first_fmt = "{levels} ({counts} penguins)") %>% 
  levels()
```

Behind the scenes, `fct_add_counts()` uses the function `fct_glue_labels()` to
construct labels using a [glue]-templating string. Therefore, `fct_glue_labels()`
would be a more appropriate function for generic relabeling using glue:

```{r}
random_penguins$species %>% 
  fct_glue_labels(
    fmt = "{tolower(levels)}", 
    first_fmt = "Species: {tolower(levels)}"
  ) %>% 
  levels()
```


#### Jekyll helpers

I also include functions I use to create and maintain my website.
`jekyll_create_rmd_draft()` creates a post in the `_R/_drafts` folder.

```{r}
withr::with_dir(tempdir(), {
  dir.create("_R")
  dir.create("_R/_drafts")
  
  # Basic use
  jekyll_create_rmd_draft(slug = "today-i-learned")
  
  # Accepts a date
  jekyll_create_rmd_draft(
    slug = "yesterday-i-learned", 
    date = Sys.Date() - 1
  )

  # Filler text used if slug is not provided
  jekyll_create_rmd_draft()
})
```



## More involved demos

These are things that I would have used in the demo above but cut and moved
down here to keep that overview succinct.

### Comparing pairs of values over a posterior distribution

I wrote `compare_pairs()` to compute posterior differences in Bayesian models.
For the sake of example, let's fit a Bayesian model of average bill length for
each species in `penguins`. We could get these estimates more directly using the
default dummy-coding of factors, but let's ignore that for now.

```{r, results = "hide"}
library(rstanarm)
m <- stan_glm(
  bill_length_mm ~ species - 1,
  penguins,
  family = gaussian
)
```

Now, we have a posterior distribution of species means.

```{r}
newdata <- data.frame(species = unique(penguins$species))

p_means <- posterior_linpred(m, newdata = newdata) %>%
  as.data.frame() %>%
  tibble::as_tibble() %>%
  setNames(newdata$species) %>%
  tibble::rowid_to_column("draw") %>%
  tidyr::gather(species, mean, -draw) %>%
  print()
```

For each posterior sample, we can compute pairwise differences of means with
`compare_means()`.

```{r pairs, fig.width = 4, fig.height = 2.5, fig.retina = 2}
pair_diffs <- compare_pairs(p_means, species, mean) %>%
  print()

library(ggplot2)

ggplot(pair_diffs) +
  aes(x = pair, y = value) +
  stat_summary(fun.data = median_hilow, geom = "linerange") +
  stat_summary(fun.data = median_hilow, fun.args = list(conf.int = .8),
               size = 2, geom = "linerange") +
  stat_summary(fun.y = median, size = 5, shape = 3, geom = "point") +
  labs(x = NULL, y = "Difference in posterior means") +
  coord_flip()
```

...which should look like the effect ranges in the dummy-coded models.

```{r, results = "hide"}
m2 <- update(m, bill_length_mm ~ species)
m3 <- update(
  m, 
  bill_length_mm ~ species, 
  data = penguins %>% mutate(species = forcats::fct_rev(species))
)
```

Give or take a few decimals of precision and give or take changes in signs
because of changes in who was subtracted from whom.

```{r}
# Adelie versus others
m2 %>% 
  posterior_interval(regex_pars = "species") %>% 
  round(2)

# Gentoo versus others
m3 %>% 
  rstanarm::posterior_interval(regex_pars = "species") %>% 
  round(2)

# differences from compare_pairs()
pair_diffs %>% 
  tidyr::spread(pair, value) %>% 
  select(-draw) %>% 
  as.matrix() %>% 
  posterior_interval() %>% 
  round(2)
```

[glue]: https://glue.tidyverse.org/index.html "glue: Interpreted String Literals"